---
title: "Soft Sediment Summaries"
author: "Ben Millard-Martin"
date: '`r date()`'
output:
  html_document:
    theme: cosmo
    code_folding: hide
    toc: true
    toc_float: true
    echo: true
    collapse: true
---

# Setup

These blocks of code sets base for the subsequent analyses:  
- loads all necessary packages   
- set working directory  
- read in diversity abundance, diversity length, ysi and habitat  
- create wide data for Vegan package  
- recreate long data with 0 observations  
- clean ysi data(remove duplicates)  
- create wide temperature data from ysi data   
- merge temperature and habitat data with diversity abundance  
- assign levels and names to month objects  
Set working directory
Read in data
Reformat data  
- create wide data  
- clean YSI data  
- extract temperature data  
- merge temperature and habitat data with diversity abundance data  
- create new month labels
 Load Packages
```{r, message=FALSE, warning= FALSE}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/bivalve_base.R")
```

# Overview
This document provides a summary of soft sediment sampling completed on Calvert Island from 2016-2017. It contains summaries of the bivalves, benthic invertebrates(meiofauna), macrophytes and sediment chlorophyll content at six sites in Pruth Bay. A summary of seasonal temperature is also provided here. Seasonal sites and protocols changed between 2016 and 2017. 

A site sampling event comprised of random sample locations within the site boundaries. At these sample locations macrophyte diversity and cover, invertebrate diversity and abundance and bivalve diversity, abundance and size data were recorded. 

In 2016 seasonal sites were Clam Garden Reference(hcg), Marten Beach(marten) and Piles Beach(piles). In 2017 Clam Garden Reference was replaced with Pruth Bay East(pbe) due to difficulties in controlling sampling effort. The 2016 methods used five 0.5m x 0.5m PVC quadrats for bivalves and were excavated to a depth of 20cm for each site sampling event.The 2017 methods used 15 0.25m x 0.25m steel box quadrats for bivalves and were excavated to a depth of 10cm for each site sampling event. For further details on the methods please see protocol documents.  

Due to the diffenences in methodology between years, count data cannot be directly compared because of differences in effort. Bivalve abundance is therefore presented as densities rather than counts per quadrat or site visit. These summaries also provide insight into diffences between methods and the potential limitations of either method. 

Sites were visited monthly or bimonthly(every two months) from June to September with winter sampling events in November and February. 

# Bivalves {.tabset}
Bivalve diversity, abundance and size data are presented in several different formats. Abundance is corrected for differences in sampling effort between years. All bivalves were identified to species aside from _Macoma spp._ less than 15mm. _Macoma spp._ less than 15mm are identified as unknown Macoma. 

## Diversity {.tabset}

```{r, message=FALSE, warning= FALSE,, results="hide", fig.show="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/bivalve_diversity.R")
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/bivalve_NMDS.R")
```

### Species Accumulation
```{r, message=FALSE, warning= FALSE}
bivalve_species_accumulation
```
Bivalve species accumulation curve. A bit of a strange curve, but we have reached the asymptote. We are capturing the entire bivalve community withing this habitat type (protected intertidal soft sediment).

### Community NMDS {.tabset}
```{r, message=FALSE, warning= FALSE, results="hide"}
bivalve_NMDS
```

### Seasonal Species Richness
```{r}
bivalve_seasonal_species_richness
```
Seasonal species richness for 2016 and 2017.

### Seasonal Shannons H
```{r}
bivalve_seasonal_shannons
```
Seasonal diversity (Shannons H) for 2016 and 2017.

### Species Richness by tide height
```{r}
bivalve_tidal_richness
```
Species richness by tide height and year. Tide height is measured as meters above chart datum(mean low tide). Tide height was restricted to 1.85m in 2016.

### Bivalve Abundance by tide height
```{r}
bivalve_tidal_abundance
```
Bivalve abundance by tide height and year. Tide height is measured as meters above chart datum(mean low tide). Tide height was restricted to 1.85m in 2016.

## Individual Species {.tabset}

### Abundances {.tabset}
```{r}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/bivalve_individual.R")
```

#### Baltic macoma
```{r}
bama
```

#### Bent-nose macoma 
```{r}
beno
```

#### Butter clam
```{r}
bucl
```

#### Nuttall's cockle
```{r}
cocl
```

#### Japanese littleneck clam
```{r}
jlne
```

#### Kennerley's venus
```{r, message=FALSE, warning= FALSE}
keve
```

#### Softshell clam
```{r}
myac
```

#### Pacific littleneck clam
```{r}
plne
```

#### Pointed macoma
```{r}
pmac
```

#### Juvenile macoma
Pointend and/or bent-nose macoma.
```{r}
umac
```

### Length {.tabset}
Focus on second figure. I can figure out how to hide first figure.

```{r, results="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/bivalve_length.R")
```

#### Baltic macoma
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("bama", "")
```

#### Bent-nose macoma 
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("beno", "")
```

#### Butter clam
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("bucl", "")
```

#### Nuttall's cockle
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("cocl", "")
```

#### Japanese littleneck clam
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("jlne", "")
```

#### Kennerley's venus
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("keve", "")
```

#### Softshell clam
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("myac", "")
```

#### Pacific littleneck clam
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("plne", "")
```

#### Pointed macoma
```{r, message=FALSE, warning= FALSE, results= "hide",fig.height = 7, fig.width = 15}
lfq_year_month("pmac", "umac")
```

#### Juvenile macoma growth
```{r}
mean_length_seasonal("umac",2017)
```

# Meiofauna (benthic inverts) {.tabset}
Invertebrate diversity and abundance are presented in several different formats. Most invertebrates were identified to higher levels of classification.

```{r, message=FALSE, warning= FALSE}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/invert_base.R")
```

## Diversity {.tabset}

```{r, message=FALSE, warning= FALSE, fig.show="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/invert_diversity.R")
```

### Species Accumulation Curve
```{r, message=FALSE, warning= FALSE}
invert_species_accumulation
```

### Seasonal Species Richness
```{r, message=FALSE, warning= FALSE}
invert_seasonal_species_richness
```

### Seasonal Shannons H
```{r, message=FALSE, warning= FALSE}
invert_seasonal_shannons
```

### Species Richness by tide height
```{r, message=FALSE, warning= FALSE}
invert_tidal_richness
```

### Invertebrate abundance by tide height
```{r, message=FALSE, warning= FALSE}
invert_tidal_abundance
```

## Individual Species {.tabset}
These figures display 2017 seasonal abundance by tide height of some of the more common species.

### Amphipods
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("amph", "Order: Amphipoda")
```

### Threaded bittium
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("thbi", expression(italic(Bittium~eschrichtii)))
```

### Acorn Barnacle
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("bagl", expression(italic(Balanus~glandula)))
```

### Limpets
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("limp", "Order: Patellogastropoda")
```

### Mussel
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("myti", expression(italic(Mytilus~spp.)))
```

### Polychaetes
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("poly", "Class: Polychaeta")
```

### Purple dwarf-venus
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("puve", "Nutricola~tantilla")
```

### Roundworms
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("rowo", "Phylum: Nematoda")
```

### Checkered periwinkle
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("scut", expression(italic(Littorina~scutulata)))
```

### Flatworms
```{r, message=FALSE, warning= FALSE}
invert_abundance_seasonal("flat", "Phylum: Platyhelminthes")
```

# Macrophytes {.tabset}
These figures present seasonal percent cover of individual species and all macrophytes combined. 

```{r, message=FALSE, warning= FALSE, fig.show="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/macro_base.R")
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/macro_cover.R")
```

## Sea Lettuce
```{r, message=FALSE, warning= FALSE}
mean_cover_seasonal("ulva", expression(italic(Ulva~lactuca)))
```

## Gutweed
```{r, message=FALSE, warning= FALSE}
mean_cover_seasonal("ulin", expression(italic(Ulva~intestinalis)))
```

## Gracilaria
```{r, message=FALSE, warning= FALSE}
mean_cover_seasonal("gras", expression(italic(Gracilaria~spp.)))
```

## Seagrass
```{r, message=FALSE, warning= FALSE}
mean_cover_seasonal("zost", expression(italic(Zostera~marina)))
```

## Total Cover
```{r, message=FALSE, warning= FALSE}
ma_total_cover
```

# Temperature {.tabset}
```{r, message=FALSE, warning= FALSE, fig.show="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/temp20152017.R")
```
These figures summarise temperaturde data collected by ibuttons(EIRP Lab) and hobo loggers(Hakai Nearshore) for 2015-2017. There are several months of missing data in 2016 when neither programs had loggers installed.

These loggers are installed in the intertidal zone at 2.2m(mlt) and have been trimmed of their "dry times".

Temperature at depth intervals is also presented. This was achieved by selecing data at specific tide heights. This informs about mixing at these sites.

## Monthly Temperature
```{r, message=FALSE, warning= FALSE}
temp_monthly
```

## Marten: temperature by depth
```{r, message=FALSE, warning= FALSE}
temp_Marten
```
Temperature at depth intervals. 

## Pruth: temperature by depth
```{r, message=FALSE, warning= FALSE}
temp_Pruth
```

## Piles: temperature by depth
```{r, message=FALSE, warning= FALSE}
temp_Piles
```

## 
```{r, message=FALSE, warning= FALSE}

```

# Chlorophyll {.tabset}
Sediment chlorophyll a concentrations were measured at monthly intervals along with seasonal sampling events. 

```{r, message=FALSE, warning= FALSE, fig.show="hide"}
source("C:/Users/FABS/Desktop/Rwd_Ben/Soft_sed/working_files/Rscript/Soft_sediment/chlorophyll.R")
```

## All sites
```{r, message=FALSE, warning= FALSE, fig.show="hide"}
chla
```
Chlorophyll A concentrations measured in nanomoles per gram sediment. All sites combined for 2016-2017.

## All sites
```{r, message=FALSE, warning= FALSE, fig.show="hide"}
chla_bysite
```
Chlorophyll A concentrations measured in nanomoles per gram sediment. All sites for 2016-2017.










